---
title: "PA1_template.Rmd"
date: "October 14, 2015"
output: html_document
---

Reproducible Research

1) Loading and processing the data
  a) Load the data (i.e. read.csv())
  b) Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
if(!file.exists("activity.csv")){
  unzip ("activity.zip")
}

activitydata <- read.csv("activity.csv", colClasses = c("integer","Date","factor"))
omitNA <- na.omit(activitydata)
rownames(omitNA) <- 1:nrow(omitNA)
head(activitydata)
head(omitNA)
dim(omitNA)
dim(activitydata)
summary(omitNA)
summary(activitydata)
```

2) What is mean total number of steps taken per day?
  a) Calculate the total number of steps taken per day
  b) Make a histogram of the total number of steps taken each day
  c) Calculate and report the mean and median of the total number of steps taken per day
```{r}
library(ggplot2)
q <- qplot(date, weight = activitydata$steps, data = activitydata, geom = "histogram")
print(q)

mean(tapply(activitydata$steps, activitydata$date, sum, na.rm = TRUE))
median(tapply(activitydata$steps, activitydata$date, sum, na.rm = TRUE))

```

3) What is the average daily activity pattern?
   a) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
   b) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
avgStepsPerInterval <- data.frame(cbind(activitydata$interval, tapply(activitydata$steps, 
    activitydata$interval, mean, na.rm = TRUE)))
colnames(avgStepsPerInterval) <- c("interval", "steps")
q <- ggplot(data = avgStepsPerInterval, aes(x = interval, y = steps)) + geom_line()
print(q)

maxTime <- avgStepsPerInterval[which.max(avgStepsPerInterval$steps), "interval"]
strftime(as.POSIXct(Sys.Date()) + as.difftime(round(maxTime/100), units = "hours") + as.difftime(maxTime%%100, units = "mins"), "%r", tz = "UTC")
```

4) Impute missing values

  a) Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

  b) Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.Create a new dataset that is equal to the original dataset but with the missing data filled in.

  c) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
  
```{r}
countNA <- sum(is.na(activitydata$steps))
print(countNA)
fillNA <- activitydata
head(fillNA)
fillNA[is.na(fillNA[, 1]), 1] <- avgStepsPerInterval[is.na(fillNA[, 1]), 2]
head(fillNA)
q <- qplot(date, weight = fillNA$steps, data = fillNA, geom = "histogram")
print(q)
mean(tapply(fillNA$steps, fillNA$date, sum))
median(tapply(fillNA$steps, fillNA$date, sum))
```

5) Are there differences in activity patterns between weekdays and weekends?
  a) Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
  b) Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
  
```{r}
fillNA$day <- as.factor(ifelse(weekdays(fillNA$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
head(fillNA)
library(lattice)
xyplot(steps ~ interval | day, aggregate(steps ~ interval + day, fillNA, FUN = mean), layout = c(1, 2), type = "l", group = day)
xyplot(steps ~ interval | day, aggregate(steps ~ interval + day, fillNA, FUN = mean), layout = c(1, 2), type = "l", ylab = 'Average No of Steps')
```